Quantum Magic Games

### 2023-07-03 paper/main.tex
paper => proof main result, first direction

\usepackage[table]{xcolor}
\usepackage{subcaption}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{nicefrac}
\usepackage{hhline}
\usepackage{bigdelim}
\usepackage{booktabs}

@@

\printbibliography{}
\end{document}

>>>>

\usepackage[parfill]{parskip} % noindent + space between paragraph (I really prefer...)
\usepackage[table]{xcolor}
\usepackage{subcaption}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{nicefrac}
\usepackage{hhline}
\usepackage{bigdelim}
\usepackage{booktabs}

@@

\begin{definition}[Parity Telepathy Game~\cite{arkhipov2012}]\label{def:parity-game}
  The parity telepathy game on an arrangement \(A = (V, E, \ell)\) is
  a two-players cooperative game (say, played by Alice and Bob). The
  game is conducted by a referee charlie. Alice and Bob may agree on a
  prior strategy, but cannot communicate once the game starts. At each
  round,
  \begin{enumerate}
  \item Charlie randomly picks a vertex \(u \in V\) and one of the hyperedges
    \(e \in E\) that contains \(u\);

  \item Charlie sends the vertex \(u\) to Alice and the hyperedge \(e\) to Bob;

  \item Alice labels the vertex \(u\) with \(a(u) = \pm{}1\) and sends
    \(a(u)\) to Charlie;

  \item Bob labels each vertex \(v \in e\) with \(b(v) = \pm{}1\) and sends
    each \(b(v)\) to Charlie;

  \item Alice and Bob win if \(a(u) = b(u)\) and \(\prod_{v} b(v) = \ell(e)\).
  \end{enumerate}
\end{definition}

Notice that the generalization above differs a bit from the magic
square game seen in Section~\ref{sec:magic-square}, where Alice
receives a hyperedge from Charlie. This modification generalizes to
arrangements that do not share the property exhibited by the magic
square that its lines can be split into two disjoint sets (rows and
columns). In this sense, one can simply understand the magic square
game as a combination of two rounds. In the first ``half'' of the
round, Charlie sends the vertex \(u\) to Alice and the line \(e\) to
Bob. In the second half, Charlie sends \(u\) to Bob, but
\emph{another} line \(e' \neq e\) to Alice.

There exists a strategy that wins with certainty a parity telepathy
game played on a magic arrangement that has a quantum realization in
which all the operators have all real eigenvalues, provided that Alice
and Bob can share entanglement in advance~\cite[Theorem
17]{arkhipov2012}. It suffices for them to share a maximally entangled
state
\[
  \rho_{AB} = \sum_{i = 0}^{n-1} \ket{i}\ket{i},
\]
where \(\{\ket{i}\}\) is a basis for the finite-dimensional
Hilbert-space \(\mathcal{H}\) of dimension \(n\). In the case of the
magic square, \(n = 4\) and
\(\{\ket{i}\} = \{\ket{0}, \ket{1}, \ket{2}, \ket{3}\}\), which,
rewritten as binary strings, is
\(\{\ket{00}, \ket{01}, \ket{10}, \ket{11}\}\). The proof is similar
to our proof for Theorem~\ref{theo:quantum-wins}, using the
simultaneous diagonalization of operators in the same hyperedge and
the symmetry of the state.

\subsection{Main Result: Characterization of Magic Arrangements by Planar Graphs}

We now move on to the main result of~\cite{arkhipov2012}. We shall
derive a characterization of magic arrangements by the planarity of
its corresponding intersection graph.

\begin{definition}[Intersection Graph]\label{def:intersec}
  The intersection graph of an arrangement \(A = (V, E)\) is the
  undirected graph \(G = (V', E')\), where \(V' = E\), and there is an
  edge between two vertices \(e_{0}, e_{1} \in V'\) for each vertex in
  the insection of \(e_{0}\) and \(e_{1}\).
\end{definition}

The intersection graph of an arrangement is a dual representation in
which the roles of vertices and hyperedges are switched. The vertices
of the intersection graph are labelled with the same values of the
corresponding hyperedges in the arrangement. Similarly, a quantum
realization of the intersection graph assigns to its edges the same
measurements ascribed to the corresponding vertices in the
arrangement. The properties observed in the arrangement are also
carried to the intersection graph. In particular, the observables
assigned to the edges around a vertex in the intersection graph (by
around, we mean the edges that touch that vertex) mutually commute and
its product equals the identity times the label of the vertex. To
illustrate, the intersection graph of the magic square is the complete
bipartite graph on six vertices \(K_{3, 3}\) and the intersection
graph of the magic pentagram is the complete graph on five vertices
\(K_{5}\).

We are also going to need the definition of a topological minor:

\begin{definition}[Topological Minor]\label{def:topo}
  A graph \(H\) is a topological minor of \(G\) if there is an
  embedding of \(H\) in \(G\) that consists of an injective map
  \(\phi\) from every vertex \(v \in H\) to a vertex
  \(\phi(v) \in G\), and a map from each edge \((u, v) \in H\) to a
  simple path from \(\phi(u)\) to \(\phi(v)\) in G, such that these
  paths are disjoint (except on their endpoints).
\end{definition}

We now restate the main result of~\cite{arkhipov2012} and its proof,
further illustrating it with examples.

\begin{theorem}\label{theo:main}
  An arrangement is magic if and only if its intersection graph is not
  planar.
\end{theorem}

We start by considering the direction that says that ``nonplanar
implies magic'', or equivalently, if the intersection graph is planar,
then its arrangement is not magic.

\begin{proof}
  First of all, assume that there exists a quantum realization of the
  arrangement (if there is none, then the arrangement is not magic).
  The goal is to show that, if there exists a quantum realization and
  the intersection graph is planar, then the parity of the arrangement
  is \(+1\) and, consequently, it is not magic. The strategy consists
  of repeatedly contracting edges of the intersection, and show that
  contraction preserves a few properties. To contract an edge is to
  merge the two endpoints \(v\) and \(w\) into a new vertex
  \(u\). Every edge that used to touch \(v\), now touches \(u\), and
  every edge that used to touch \(w\), now touches
  \(u\). Consequently, the edge that was contracted becomes two
  self-loops. Moreover, we define the label of the new vertex \(w\) as
  the product of the labels of \(u\) and \(v\). We observe that
  contraction preserves the following properties:

  \begin{enumerate}
  \item The resulting graph remains planar.

  \item The parity (product of all the vertex labels) remains the
    same. This follows directly from the definition of the label of
    thenew vertex.

  \item The product of the labels of edges around any vertex, in
    cyclic order (say, without loss of generality, counterclockwise),
    equals the identity times the label of the vertex. To see this,
    first notice that this property holds for the initial graph, since
    all observables around a vertex pairwise commute (recall that
    these observables correspond to the observables assigned by the
    quantum realization to vertices that lie in the hyperedge in the
    arrangement). However, the commutative property does not
    necessarily hold after contraction. Nevertheless, if we follow a
    cyclic order, we can still see that this property is
    preserved. Let \(M_{0}, \ldots, M_{n-1}\) be the observables
    around a vertex \(v\), going counterclockwise. If
    \(M_{0} \ldots M_{n - 1} = \ell(v) I\), then for any starting
    point \(k\),
    \(M_{k} M_{k+1}\ldots M_{n-1}M_{0}\ldots M_{k-1} = \ell(v) I\),
    since
    \begin{align}
      M_{k}M_{k+1} \ldots M_{n-1}M_{0}\ldots M_{k-1}
      &= (M_{k-1}\ldots M_{0})(M_{0} \ldots M_{n-1})(M_{0}\ldots M_{k-1})  \label{eq:cyclic1} \\
      &= (M_{k-1}\ldots M_{0})\ell(v) I(M_{0}\ldots M_{k-1}) = \ell(v) I,  \label{eq:cyclic2}
    \end{align}
    where (\ref{eq:cyclic1}) follows from the fact that
    \(M_{i} M_{i} = \ell(v) I\). Now, let \(X\) be the measurement assigned to
    the edge that is being contracted and let the labels around its
    endpoints \(v\) and \(w\) be \(M_{0}\ldots M_{m-1}X\) and
    \(X N_{0}\ldots N_{n-1}\) going counterclockwise. Then,
    \(M_{0} \ldots M_{m-1} X = \ell(v)I\) and
    \(X N_{0} \ldots N_{n-1} = \ell(w)I\). Multiplying these gives
    \[
      M_{0}\ldots M_{m-1} X^{2} N_{0} \ldots N_{n-1} =
      M_{0}\ldots M_{m-1} N_{0} \ldots N_{n-1} =
      \ell(v) \ell(w) I = \ell(u) I,
    \]
    so the product of the observables around the new vertex \(w\)
    still is the identity times its label. Notice that key to this
    strategy is the fact that the graph is planar, and thus we can
    draw the edges around the new vertex without mixing the edges that
    used to touch \(v\) and \(w\).
  \end{enumerate}

  By repeatedly contracting edges, eventually we reach a graph with a
  single vertex.  Due to how we defined the label of each new vertex
  (i.e.\ as the product of the labels of the vertices being merged),
  it follows that the label of the last vertex equals the product of
  the labels of all the original vertices, and therefore equals the
  parity of the arrangement. The remaining vertex has multiple
  self-loops, two for each measurement \(M\) ascribed to the original
  edges.  Since \(M^{2} = I\), going counterclockwise we get a
  sequence \(M_{0}^{2} \ldots M_{\vert E'\vert -1}^{2} = I\). But, we
  saw that contraction preserves the third property, which says that
  the observables around a vertex multiply to the identity times the
  vertex label. Thus, the label of the final vertex is \(1\) and,
  consequently, the parity of the arrangement is \(+1\). \qed
\end{proof}
\printbibliography{}
\end{document}
